The first high-redshift Changing Look Quasars

### 2019-11-29 tex/CIV_CLQs.tex
major, major update

\date{Accepted XXX. Received YYY; in original form ZZZ}
\pubyear{2018}

@@

\begin{abstract}
We report on three redshift $z>2$ quasars with dramatic changes in
their \civ\ emission lines, the first ``Changing-Look'' quasars at high
redshift.  This is also the first time the changing-look behaviour has
been seen in a high-ionization emission line.
%%
SDSS J1205+3422, J1638+2827 and J2228+2201 show interesting behaviour
in their observed optical light curves, and subsequent spectroscopy
shows significant changes in the \civ\ broad emission line, with both
line collapse and emergence being displayed in rest-frame timescales
of $\sim$240-1640 days.
%%
Where observed, the profile of the Ly$\alpha$/\nv emission complex
also changes, and there is tentative evidence for changes in the \mgii
line.
%%
Although line measurements from the three quasars show large changes
in the \civ\ flux-line width plane, the quasars are not seen to
be outliers when considered against the full $z\sim2$ quasar population
in terms of (rest) Equivalent Width and FWHM properties.
%%
We put these observations in context with recent ``state-change''
models, but note that even in their `low-state', the \civ\ CLQs are
above $\sim$10\% in Eddington luminosity.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
accretion, accretion discs -- surveys -- quasars: general -- quasars: individual: J1100-0053 
\end{keywords}

@@

The optical continuum variability of quasars has been recognized since
their first optical identification
\citep[e.g.,][]{MatthewsSandage1963, MacLeod2012}.  Dramatic changes
in the broad emission lines (BELs) of quasars has only recently been
identified \citep[e.g., ][]{LaMassa2015}.  Samples of over 100
``Changing Look'' quasars (CLQs) or ``Changing State'' quasars (CSQs)
have now been assembled \citep[e.g.][]{MacLeod2019, Graham2019b}. The
community uses both these terms as a cover for the underlying
physics. For sake of argument, CLQs can potentially be thought of as
the extension to the BELs of quasar continuum variability \citep[e.g.,
][]{MacLeod2012} whereas the CSQs have a `state-transition' similar to
that in Galactic X-ray binaries \citep[][]{NodaDone2018, Ruan2019}. In
this paper, we use the term `Changing Look', as we are currently
agnostic, and confessedly ignorant, to the underlying physical
processes.

@@

%                              &    / \AA\    &      / eV             &  / eV            &   Lower                  & Conf.      &  $J$       & Upper                      &  Conf.                       & $J$       & / cm$^{-1}$    &  / s$^{-1}$                \\      \hline

@@

    \caption{Strong UV/optical spectral emission lines in quasars, and their atomic data.
      Data from the NIST Atomic Spectra Database       \citep{Kramida2018, Kramida2019}.
      The Transition Energies are $E=hc/\lambda$ for the given wavelength. The Ionization Energy is the
      energy required to ionize the given species, e.g. 64.49 eV are needed to create a \cv\ ion.
      Transitions Level configurations are given in standard spectroscopic notation
      %($N^{2s+1}$ $L_{j}$. 
      $A_{i,j}$ are transition probabilities.
%      $A_{i,j}$ are \href{https://www.nist.gov/pml/atomic-spectroscopy-compendium-basic-ideas-notation-data-and-formulas/atomic-spectroscopy-6}{transition probabilities}.} 
%      such that the radiative lifetime $\tau_{k}$ of an atomic level $k$ is related to the sum of transition probabilities to all levels $i$ lower in energy than $k$
%      The 1548.202 and 1550.774 \AA\ emission doublet is created by the 2$p$ $^{2}P_{0}$ $-$ 2$s$ $^{2}S$ transition, split by total angular momentum $J= 3/2$ and $1/2$. 
      Data from the NIST Atomic Spectra Database       \citep{Kramida2018, Kramida2019}.}
    %%  Mg II Z = 12, Na isoelectronic sequence  

@@

While there have been many studies on triply ionized carbon, i.e
\civ, these have tended to focus on broad absorption line quasars
\citep[BAL QSOs; see Table 1 of][]{Hemler2019} or the Baldwin Effect
\citep[BEff; ][]{Baldwin1977, Bian2012, Jensen2016,
Hamann2017}\footnote{As noted in \citet{Rakic2017}, two different
types of Baldwin effect are present in the literature: the {\it
global} (or {\it ensemble}) Baldwin Effect, which is an
anti-correlation between the Equivalent Width (EW) of the emission line and the
underlying continuum luminosity of {\it single-epoch} observations of
a {\it large number} of AGN and second, the {\it intrinsic} Baldwin
effect, the same anti-correlation but in an {\it individual, variable}
AGN \citep{PoggePeterson1992}.}.  Dramatic changes in the
collisionally excited broad {\it emission} line (BEL) of \civ\ and
indeed \ciii have not to this point been reported.

\begin{table*}
  \centering
  \begin{tabular}{l l   r ll  r r l}
    \hline 
    \hline 
    \multirow{2}{*}{Object} & \multirow{2}{*}{Redshift} &  $g$-band     & \multirow{2}{*}{MJD} &  \multirow{3}{*}{Instrument}  & Exposure     &   SDSS               & \multirow{2}{*}{Notes} \\
                            &                                                     &   (mag)           &                                 &                                               &  Time (secs) &   Plate-FiberID  & \\
    \hline  
    J120544.7+342252.4   & 2.068                               &   18.27             &  53498                   & SDSS                                    &  8057            & 2089-427             & \\
                                        & 2.071                                &                        &  58538                   & DBSP                                     &  1800            &  ---                      &  Average conditions \\
                                         & 2.071                               &                         &  58693                  & DBSP                                     &  2400            &  ---                      &          \\
                                         &                                       &                          &                               &                                              &                      &                               &                               \\
    J163852.9+282707.7   & 2.185                            &   19.77               &  54553                   & SDSS                                      &   4801            &  2948-614              & \\
    	                                 &  2.186                            &                          &  55832                  & BOSS                                      &   3600            &  5201-178            & \\
                                         &  2.182                           &                          &  58583                     & LRIS                                     &  1800              &  ---                      & \\
                                        &                                      &                          &                                 &                                              &                      &                              &                                 \\
    J222818.7+220102.9   & 2.217                           & 19.97                &  56189                  &  BOSS                                       &  2700            &   6118-720          & \\
                                          & 2.222                         &                          &  56960                    & BOSS                                       &  4500            &   7582-790          & eBOSS reobservation \\ 
                                         &  2.222                         &                          &  58693                  &  DBSP                                         & 2400             &    ---                        &    \\
    \hline \hline   
  \end{tabular}
  \caption{
    Details of our spectroscopic observations.
%    Redshift and redshift errors from SDSS SkyServer for SDSS, BOSS and eBOSS spectra.
    Redshift errors are typically  $\pm$0.002. 
    Exposure times are from the {\tt plate.fits} file.  SDSS, BOSS and  eBOSS spectra have $\mathcal{R}\sim$2,000.
    DBSP: Double Spectrograph on the Palomar 200-inch telescope.
    LRIS:  Low Resolution Imaging Spectrometer on Keck I 10m telescope.
    %  2019-Feb-24    & DBSP            &  2$\times$900 f
    %  2019-Jul-29     & DBSP            &  2$\times$1200
    %  2019-Apr-10     & LRIS              &  2$\times$900
    %& 2019-Jul-29       & DBSP             & 2$\times$1200
  } 
  \label{tab:obs_notes}
\end{table*}

@@

This paper is organised as follows. In Section 2, we describe our
sample selection, catalogues, and observational data sets.  In Section 3,
we show the high-$z$ quasars and report the line properties for the
quasars at the observed epochs.  We give a very brief theoretical
discussion in Section 4. We present our conclusions in Section 5.  We
report all magnitudes on the AB zero-point system \citep{Oke_Gunn1983,
Fukugita1996} unless otherwise stated. For the WISE bands,
$m_{\rm AB} = m_{\rm Vega} + m$ where $m = (2.699, 3.339)$ for WISE W1
at 3.4$\mu$m and WISE W2 at 4.6$\mu$m, respectively
\citep{Cutri2011}.
We adopt a flat $\Lambda$CDM cosmology with $\Omega_{\Lambda} = 0.73$,
$\Omega_{\rm M} = 0.27$, and $h = 0.71$.
% in order to be consistent with \citet{Hamann2017}. As a guide this cosmology has a $z=2.000$ comoving radial distance of 5244.3 Mpc, a $\sim$1.25\% difference compared to 5179.0 Mpc from $\Omega_{\Lambda} = 0.70$, $\Omega_{\rm M} = 0.30$, and $h = 0.70$ that is used in \citet{Shen2011}.

@@

\subsection{Multi-Wavelength Properties}
Mid-infrared data (3.4 and 4.6$\mu$m) is available from the beginning
of the {\it Wide-field Infrared Survey Explorer} ({\it WISE}) mission
\citep[2010 January; ][]{Wright2010} through the fifth-year of {\it
NEOWISE-R} operations \citep[2018 December; ][]{Mainzer2011}. The {\it
WISE}WISE scan pattern leads to coverage of the full-sky approximately
once every six months (a ``sky pass''), but the satellite was placed
in hibernation in 2011 February and then reactivated in 2013
October. Hence, our light curves have a cadence of 6 months with a
32-month sampling gap. [{\it DANIEL to write something on FIRST
non-detections; possibly check the X-rays too?}]

\subsection{CLQ Selection}
{\bf MJG to finalise this} \\
Our high-$z$ CLQs were identified as follows.  We selected all 64,774
SDSS DR15 sources with $z > 0.35$ classified as `QSO', having at least
two spectra separated by at least 100 days, and with a corresponding
CRTS light curve. We fitted a damped random walk to the CRTS data via
Gaussian process regression and the photometric magnitudes at the
epochs of the SDSS spectra for a given source are predicted. Those
where $|\Delta V| > 0.3$ are then selected for visual
inspection. Only three quasars SDSS J120544.7+342252.4 (hereafter
J1205+3422), SDSS J163852.93+282707.7 (hereafter J1638+2827) and SDSS
J222818.76+220102.9 (hereafter J2228+2201), satisfied these selection
criteria and showed interesting or dramatic emission line behaviour.

@@

\subsubsection{Spectrophotometry}
{\bf MJG to finish this} \\
We want all our spectra to have reliable flux calibrations.  SDSS
spectra ae spectrophotometrically calibrated.  BOSS and eBOSS spectra
have spectrophotometric corrections applied in the latest data release
\citep{Hutchinson2016, Jensen2016, Margala2016}.  Due to the high-$z$
of our objects, \oiii is not available to us to use as a calibrating
flux line. Instead we use photometric data from the ZTF since all our
non-SDSS/BOSS/eBOSS data are taken after MJD 57500.

@@

In particular we use the Quasar Spectral Fitting (QSFit) software
package presented in \citet{Calderone2017}. This provides luminosity
estimates as well as width, velocity offset and equivalent width of 20
emission lines, including \civ, \ciii and \mgii.  We process and fit
all nine spectra using the lastest version (v1.3.0) of the QSFit 
\href{https://qsfit.inaf.it/cat_1.30/onlinefit.php}{online
calculator}.
% and setting $E(B-V)=0.00$.
%%
The host galaxy and blended iron emission at rest-frame optical
wavelengths components are automatically disabled when they can not be
constrained by the available data, such as the case for all our
objects (we do not have infrared spectral data).
%%
Power-law continuum slopes, $\alpha$, where $f_{\lambda} \propto
\lambda^{\alpha}$, are also reported in these catalogues and from
QSFit.  We quote the most appropriate value given the emission line
wavelength.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\subsection{Photometric and Spectral Evolution}
Figure~\ref{fig:civ_clqs} presents the optical and mid-IR light
curves for three high-$z$ CLQ quasars.  Figure~\ref{fig:civ_clqs} also
shows the spectra for each epoch, with the MJD of observation given by
the dashed vertical lines in the light curves.

@@

\begin{figure}
  \centering
  \includegraphics[width=8.7cm, trim=0.2cm 0.2cm 0.2cm 0.2cm, clip]
  {figures/Wilhite_2006_Fig2_redux_20190926.png}
   \vspace{-12pt}
   \caption[]{ {\bf NPR to update this}. 
     The change in \civ\ line width vs. line flux change. 
We compare our object J1638+2827 with the sample 
from \citet{Wilhite2006} with a sample of 105 quasars observed at
multiple epochs by the SDSS. J1638+2827 is a substantial outlier 
in this parameter space.}
  \label{fig:Wilhite2006_comparison}
\end{figure}
%\subsection{Emission Line Evolution}
Emission line measurements from the QSFit routine of
\citet{Calderone2017} are presented in Table~\ref{tab:LineMeasurements}. 
and fits for the \civ\ line are given in~\ref{fig:CIV_LineFits}.

From Figure~\ref{fig:Wilhite2006_comparison} there appears to be a
strong correlation between the change in the line flux and the change
in the line width.  Figure~\ref{fig:Wilhite2006_comparison} shows the
epoch-to-epoch flux ratio versus the ratio of line widths.

\begin{figure}
  \centering
  \includegraphics[width=8.5cm, trim=0.4cm 0.4cm 0.2cm 0.4cm, clip]
  {figures/CIV_fits_20191108.png}
   \vspace{-12pt}
   \caption[]{The \civ\ emission line and the subsequent QSFit model and
     parameters.}
   \label{fig:CIV_LineFits}
\end{figure}

\begin{table*}
  \centering
  % \input{emission_line_comparison_tab}
  \input{emission_line_tab}
  \caption{Line Measurement information for the nine epochs for the 3
    quasars.  Line Luminosity is in units of 10$^{42}$ erg s$^{-1}$; FWHM
    and $V_{\rm off}$ in km s$^{-1}$, where positive values of $V_{\rm
      off}$ means the line is blueshifted.  Equivalent widths are in units
    of \AA\ .  Positive values of V$_{\rm off}$ means the line is
    blueshifted.  All measuremennts are from QSFit, except for J1638+2827
    (MJD 54553) which are from \citet{Shen2011}.  $^{(d)}$Emission line
    modelled with two components.}
 \label{tab:LineMeasurements}
\end{table*}

\begin{table}
  \begin{centering}
    \begin{tabular}{l c l l  ll}
      \hline
      \hline
        Object       &  MJD     & \multicolumn{2}{c}{1450\AA\ Cont. }     & \multicolumn{2}{c}{3000\AA\  Cont.}  \\
                         &             &  Lumin.          &  slope                              &  Lumin.  &    slope      \\
      \hline
                         & 53498   &  40040$\pm$34    & -1.66$\pm$0.01    & ---        & ---  \\
 J1205               & 58538   &   ---                     &  ---                         & ---        & ---   \\
                         & 58693   &   ---                     &  ---                         & ---        & --- \\
                         &              &                              &                                  &       & \\
                         & 54553   &  3774$\pm$111   &   -1.47$\pm$0.20    & ---                      & --- \\
 J1638              & 55832    &  2433$\pm$14     &  -2.32$\pm$0.05    & 958$\pm$29        & -2.27$\pm$0.05 \\  
                         & 58583   &  3292$\pm$9       &    -0.54$\pm$0.01  & 4595$\pm$18      &  -0.54$\pm$0.01 \\
                         &              &                              &                                 &       &     \\ 
                         & 56189    &  ---                     &  ---                         & 1114$\pm$107    & -0.46$\pm$0.21     \\
J2228                & 56960   &  7889$\pm$14     &  -1.73$\pm$0.02   &  4634.1$\pm$28  &  -1.72$\pm$0.02      \\
                         & 58693   &  ---                      &  ---                         &  ---                       &  ---  \\
      \hline
      \hline
    \end{tabular}
    \caption{Quasar continuum $\nu L_{\nu}$ luminosities and slopes, at
      the fixed rest-frame wavelengths 1450\AA\ and 3000\AA\ from QSFit.  The
      luminosities are in units of 10$^{42}$~erg~s$^{-1}$.}
    \label{tab:Baldwin}
  \end{centering}
\end{table}
%% %% \eta_{\rm Edd} = log10((10**LBol)/((1.26e38)*(10**MBH)))
%% Using    C IV

>>>>

\date{Accepted XXX. Received YYY; in original form ZZZ}

@@

\begin{abstract}
We report on three redshift $z>2$ quasars with dramatic changes in
their \civ\ emission lines, the first ``Changing-Look'' quasars at
high redshift.  This is also the first time the changing-look
behaviour has been seen in a high-ionization emission line.  SDSS
J1205+3422, J1638+2827 and J2228+2201 show interesting behaviour in
their observed optical light curves, and subsequent spectroscopy shows
significant changes in the \civ\ broad emission line, with both line
collapse and emergence being displayed in rest-frame timescales of
$\sim$240-1640 days.  Where observed, the profile of the
Ly$\alpha$/\nv emission complex also changes, and there is tentative
evidence for changes in the \mgii line.  Although line measurements
from the three quasars show large changes in the \civ\ flux-line width
plane, the quasars are not seen to be outliers when considered against
the full $z\sim2$ quasar population in terms of (rest) Equivalent
Width and FWHM properties.  We put these observations in context with
recent ``state-change'' models, but note that even in their
`low-state', the \civ\ CLQs are above $\sim$10\% in Eddington
luminosity.
\end{abstract}

\begin{keywords}
accretion, accretion discs -- surveys -- quasars: general -- quasars: time-domain: 
\end{keywords}

@@

The optical continuum variability of quasars has been recognized since
their first optical identification
\citep[e.g.,][]{MatthewsSandage1963, MacLeod2012}.  Dramatic changes
in the broad emission lines (BELs) of quasars has only recently been
identified \citep[e.g., ][]{LaMassa2015}.  Samples of over 100
``Changing Look'' quasars (CLQs) or ``Changing State'' quasars (CSQs)
have now been assembled \citep[e.g.][]{MacLeod2019, Graham2019b}. The
community uses both these terms as a cover for the underlying
physics. For sake of argument, CLQs can potentially be thought of as
the extension to the BELs of quasar continuum variability \citep[e.g.,
][]{MacLeod2012} whereas the CSQs have a `state-transition' similar to
that in Galactic X-ray binaries \citep[][]{NodaDone2018, Ruan2019a}. In
this paper, we use the term `Changing Look', as we are currently
agnostic, and confessedly ignorant, to the underlying physical
processes.

@@

@@

    \caption{
      Strong UV/optical spectral emission lines in quasars, and their
      atomic data.  Data from the NIST Atomic Spectra Database
      \citep{Kramida2018, Kramida2019}.  The Transition Energies are
      $E=hc/\lambda$ for the given wavelength. The Ionization Energy is the
      energy required to ionize the given species, e.g. 64.49 eV are needed
      to create a \cv\ ion.  Transitions Level configurations are given in
      standard spectroscopic notation $A_{i,j}$ are transition
      probabilities.  Data from the NIST Atomic Spectra Database
      \citep{Kramida2018, Kramida2019}.}

@@

While there have been many studies on triply ionized carbon, i.e
\civ, these have tended to focus on broad absorption line quasars
\citep[BAL quasars; see Table 1 of][]{Hemler2019} or the Baldwin Effect
\citep[BEff; ][]{Baldwin1977, Bian2012, Jensen2016,
Hamann2017}\footnote{As noted in \citet{Rakic2017}, two different
types of Baldwin effect are present in the literature: the {\it
global} (or {\it ensemble}) Baldwin Effect, which is an
anti-correlation between the Equivalent Width (EW) of the emission line and the
underlying continuum luminosity of {\it single-epoch} observations of
a {\it large number} of AGN and second, the {\it intrinsic} Baldwin
effect, the same anti-correlation but in an {\it individual, variable}
AGN \citep{PoggePeterson1992}.}.  Dramatic changes in the
collisionally excited broad {\it emission} line (BEL) of \civ\ and
indeed \ciii have not to this point been reported.

\begin{table*}
  \centering
  \begin{tabular}{l l   r ll  r r l}
    \hline 
    \hline 
    \multirow{2}{*}{Object} & \multirow{2}{*}{Redshift} & $g$-band & \multirow{2}{*}{MJD} & \multirow{3}{*}{Instrument} & Exposure      &   SDSS               & \multirow{2}{*}{Notes} \\
                                         &                                        &   (mag)      &                                 &                                             &  Time (secs) &   Plate-FiberID  & \\
    \hline  
                                       & 2.068                              & 18.27          &  53498                   & SDSS                                     &  8057            & 2089-427             & \\
 J120544.7+342252.4     & 2.071                             &                    &  58538                   & DBSP                                     &  1800            &  ---                      &  Average conditions \\
                                       & 2.071                             &                     &  58693                  & DBSP                                      &  2400            &  ---                      &          \\
                                       &                                       &                     &                               &                                              &                      &                               &                               \\
                                      & 2.185                              & 19.77           &  54553                   & SDSS                                     &   4801            &  2948-614              & \\
J163852.9+282707.7     & 2.186                              &                    &  55832                     & BOSS                                    &   3600            &  5201-178            & \\
                                      & 2.182                              &                     &  58583                     & LRIS                                    &  1800              &  ---                      & \\
                                      &                                         &                    &                                 &                                            &                      &                              &                                 \\
                                      & 2.217                               & 19.97          &  56189                    &  BOSS                                    &  2700            &   6118-720          & \\
 J222818.7+220102.9   & 2.222                               &                    &  56960                     & BOSS                                    &  4500            &   7582-790          & eBOSS reobservation \\ 
                                     &  2.222                               &                    &  58693                    &  DBSP                                    & 2400             &    ---                        &    \\
    \hline \hline   
  \end{tabular}
  \caption{Details of our spectroscopic observations.  Redshift errors are
    typically $\pm$0.002.  SDSS, BOSS and eBOSS spectra have
    $\mathcal{R}\sim$2,000.  DBSP: Double Spectrograph on the Palomar
    200-inch telescope.  LRIS: Low Resolution Imaging Spectrometer on Keck
    I 10m telescope.} 
  \label{tab:obs_notes}
\end{table*}

@@

This paper is organised as follows. In Section 2, we describe our
sample selection, catalogues, and observational data sets.  In Section
3, we show the high-$z$ quasars and report the line properties for the
quasars at the observed epochs.  We give a very brief theoretical
discussion in Section 4. We present our conclusions in Section 5.  We
report all magnitudes on the AB zero-point system \citep{Oke_Gunn1983,
Fukugita1996} unless otherwise stated. For the WISE bands, $m_{\rm AB}
= m_{\rm Vega} + m$ where $m = (2.699, 3.339)$ for WISE W1 at
3.4$\mu$m and WISE W2 at 4.6$\mu$m, respectively \citep{Cutri2011}. We
adopt a flat $\Lambda$CDM cosmology with $\Omega_{\Lambda} = 0.73 $,
$\Omega_{\rm M} = 0.27$, and $h = 0.71$. All logarithms are to the base 10.

@@

\subsection{Multi-Wavelength Properties}
{\bf DS to write something on FIRST non-detections; possibly check the X-rays too?} \\
Mid-infrared data (3.4 and 4.6$\mu$m) is available from the beginning
of the {\it Wide-field Infrared Survey Explorer} ({\it WISE}) mission
\citep[2010 January; ][]{Wright2010} through the fifth-year of {\it
NEOWISE-R} operations \citep[2018 December; ][]{Mainzer2011}. The {\it
WISE}WISE scan pattern leads to coverage of the full-sky approximately
once every six months (a ``sky pass''), but the satellite was placed
in hibernation in 2011 February and then reactivated in 2013
October. Hence, our light curves have a cadence of 6 months with a
32-month sampling gap.

\subsection{CLQ Selection}
{\bf MJG to finalise this} \\
Our high-$z$ CLQs were identified as follows.  We selected all 64,774
SDSS DR15 sources with $z > 0.35$ classified as `quasar', having at least
two spectra separated by at least 100 days, and with a corresponding
CRTS light curve. We fitted a damped random walk to the CRTS data via
Gaussian process regression and the photometric magnitudes at the
epochs of the SDSS spectra for a given source are predicted. Those
where $|\Delta V| > 0.3$ are then selected for visual
inspection. Only three quasars SDSS J120544.7+342252.4 (hereafter
J1205+3422), SDSS J163852.93+282707.7 (hereafter J1638+2827) and SDSS
J222818.76+220102.9 (hereafter J2228+2201), satisfied these selection
criteria and showed interesting or dramatic emission line behaviour.

@@

\subsubsection{Spectrophotometry}
{\bf MJG to check/finalise} \\
We want all our spectra to have reliable flux calibrations.  SDSS
spectra ae spectrophotometrically calibrated.  BOSS and eBOSS spectra
have spectrophotometric corrections applied in the latest data release
\citep{Hutchinson2016, Jensen2016, Margala2016}.  Due to the high-$z$
of our objects, \oiii is not available to us to use as a calibrating
flux line. Instead we use photometric data from the ZTF since all our
non-SDSS/BOSS/eBOSS data are taken after MJD 57500.

@@

In particular we use the Quasar Spectral Fitting (QSFit) software
package presented in \citet{Calderone2017}. This provides luminosity
estimates as well as width, velocity offset and equivalent width of 20
emission lines, including \civ, \ciii and \mgii.  We process and fit
all nine spectra using the lastest version (v1.3.0) of the QSFit
\href{https://qsfit.inaf.it/cat_1.30/onlinefit.php}{online
calculator}. The host galaxy and blended iron emission at rest-frame
optical wavelengths components are automatically disabled when they
can not be constrained by the available data, such as the case for all
our objects (we do not have infrared spectral data).  Power-law
continuum slopes, $\alpha$, where $f_{\lambda} \propto
\lambda^{\alpha}$, are also reported in these catalogues and from
QSFit.  We quote the most appropriate value given the emission line
wavelength.
  \small
    \begin{tabular}{l  l | c c | c c c c | c}
              &          &  \multicolumn{2}{c}{Cont. @ 1450\AA }                  &   \multicolumn{4}{c}{CIV 1549\AA}                                                                                & Virial product \\
Object    & MJD  &      $\nu L_{\nu}$    &         Slope                             &   Luminosity         &     FWHM                 &    V$_{\rm off}$    & EW                                 & log($\nu L_{\nu}^{0.5} \times {\rm FWHM}^2$) \\
 & & $10^{42}$ erg s$^{-1}$ & (F$_\lambda \propto \lambda^\alpha$)  & $10^{42}$ erg s$^{-1}$ &  km s$^{-1}$   &     km s$^{-1}$   &  \AA                              &  $\log(M/M_{\odot})$ \\
                     & 53498   &  38129  $\pm$ 39  &  -1.57 $\pm$ 0.01 &   898 $\pm$ 15    &   6024 $\pm$  120  &  944 $\pm$  33  & 37.78 $\pm$  0.62  &   9.85 $\pm$ 0.02\\
 J1205+3422 & 58538    &  8550 $\pm$ 11    &  -1.41 $\pm$ 0.01 &  385.2 $\pm$ 3.9 &   7109 $\pm$   91   & 1085 $\pm$  29  &   71.32 $\pm$  0.71  &   9.67 $\pm$ 0.01\\  
                     & 58693$^*$ & 2725 $\pm$ 40 &  -1.27 $\pm$ 0.05 &  161 $\pm$ 22    &  14997 $\pm$ 2500 &  922 $\pm$ 690  & 91.68 $\pm$ 12.73  &  10.07 $\pm$ 0.13\\  
                      &              &                               &                               &                              &                                 &                             &                                   &                              \\
                      & 54553   & 3579 $\pm$ 49    &  -1.16 $\pm$ 0.07  & 293.6  $\pm$ 8.8 &  4630 $\pm$  180   & 183 $\pm$  57   & 127.67 $\pm$  3.80  &    9.11 $\pm$ 0.04\\  
 J1638+2827 & 55832   & 2340 $\pm$ 15    &  -2.17 $\pm$ 0.05  &  84.6  $\pm$ 4.2  &  4733 $\pm$  290   & 172 $\pm$  89   &  59.61 $\pm$  2.98  &    9.04 $\pm$ 0.05\\  
                      & 58583  & 7793 $\pm$ 19     &  -2.02 $\pm$ 0.01  & 367.5  $\pm$ 4.3 &  4511 $\pm$   71    &  94 $\pm$  24   &  77.89 $\pm$  0.91  &  9.25 $\pm$ 0.01\\  
                      &             &                               &                                &                              &                                 &                            &                                   &                            \\
                     & 56189$^*$ & 607 $\pm$ 49  &  -0.00 $\pm$ 0.16 &  72.8  $\pm$ 11.3 & 14993 $\pm$ 2400 &  828 $\pm$ 691   & 162.89 $\pm$ 25.37  &  9.74 $\pm$ 0.14\\  
 J2228+2201 & 56960   & 7842 $\pm$ 25     &  -1.72 $\pm$ 0.02 & 301.0  $\pm$ 7.2  &  7136 $\pm$  210   & -276 $\pm$  64   &  61.99 $\pm$  1.48  &    9.65 $\pm$ 0.03\\ 
              & 58693   & 2388.4 $\pm$ 6.8 &  -1.22 $\pm$ 0.01 & 145.4  $\pm$ 1.5  &  6084 $\pm$   83    &   168 $\pm$  28   &  94.77 $\pm$  0.99  &    9.26 $\pm$ 0.01\\
    \caption{Continuum at 1450\AA\ and \civ\ spectral measurements for the
      three quasar considered in this work, at all observation epochs, as
      calculated by QSFit.  The last column shows the virial product
      calculated as $\nu L_{\nu}^{0.5} \times {\rm FWHM}^2$.  $^{*}$The
      \civ\ line is very faint (with respect to the continuum), and the
      associated estimates are likely unreliable.}
    \label{tab:QSFit-results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\subsection{Photometric and Overal Spectral Evolution}
Figure~\ref{fig:civ_clqs} presents the optical and mid-IR light
curves for three high-$z$ CLQ quasars.  Figure~\ref{fig:civ_clqs} also
shows the spectra for each epoch, with the MJD of observation given by
the dashed vertical lines in the light curves.

@@

\subsection{\civ\ Emission Line Evolution}
We analyzed the spectra of the three quasars, at all observational
epochs, using the QSFit spectral fitting package
\citep{Calderone2017}.  One advantage of using QSFit is that it allows
constraints on the slope and luminosity of the broad band continuum of
the source. The relevant estimated quantities, including continuum
luminosity and slope at rest-frame 1450\AA, \civ\ line luminosity,
FWHM and EW are given in Tab.~\ref{tab:QSFit-results}. All fits are
performed with $E(B-V) = 0$ and the best fit model in the region of
the \civ\ emission line are shown in Figure~\ref{fig:QSFit-CIV}.

All \civ\ lines are fitted with a single, broad, Lorentzian profile.
This allows us to account for the narrow peak of the \civ\ line. 
No narrow components are considered for several reasons: in the epochs
of highest brightness the ``narrow'' component would have FWHM
$\sim2-3 \times 10^{3}$ km s$^{-1}$, i.e. values exceeding the usual
widths of genuine narrow lines ($\lesssim 10^3$ km s$^{-1}$); by
allowing a second component, to have such large widths their
parameters would become highly degenerate with the ``broad''
components, and the latter would also have much larger FWHM
($\sim10^{4}$ km s$^{-1}$); and by neglecting the narrow components we
have a consistent fit across all epochs.

The quasar continuum, evaluated at 1450\AA\ , and the \civ\ line
luminosities follow a similar evolution, with a ratio of $\sim$20-30,
confirming that the main driver for emission line variability is
likely the broad band continuum itself.  For all sources except
J1638+2827, the slope of the continuum changes with luminosity
following a ``bluer-when-brighter'' pattern, suggesting that a
distinct emerging component is responsible or both the slope and
luminosity variations.  In J1638+2827 the opposite behaviour is
observed, especially in the first observation epoch.  However, this
may be a bias due to the limited wavelength range available which
extends to $\lambda \sim 1240\AA$ for the first epoch, while it extend
to shorter wavelengths for the other epochs (respectively 1140\AA\ and
1010\AA).  The latter suggests that the ``emerging'' component is more
prominent at UV wavelengths, and a sufficient wavelength coverage is
required to detect it. We find there is no need for a narrow component
in J2228+2201 MJD 56189.  This is because the data resemble a P-Cygni
profile, with blueshifted absorption. This feature is very narrow, and
the uncertainties large, such that the overall $\chi^{2}$ is only
marginally affected by the addition of a further component. We
therefore avoid using a narrow component for this spectrum only, since
that would be inconsistent with the other analyses.

In all cases where the \civ\ line profile is reliably constrained the
\civ\ FWHM is approximately constant, with maximum variations
$\lesssim$~1000 km s$^{-1}$, despite the significantly larger
variations in the line luminosities.  The FWHM of broad lines is
likely related to the mass of the supermassive black hole powering the
quasar phenomenon, which is supposed to be constant on any human
timescale.  Hence it is instructive to check whether the virial
product, which is the basic quantity used to calculate the single
epoch mass estimate, show any variation.  The virial product,
calculated as $\nu L_{\nu}^{0.5} \times {\rm FWHM}^2$, is reported in
the last column of Tab.~\ref{tab:QSFit-results}.  The uncertainty
typically associated to the single epoch mass estimate is $\sim$0.5
dex, hence the virial product at all epochs are remarkably constant
and compatible with a single value of black hole mass for each source,
even in those cases where the \civ\ estimates are possibly
unreliable. The object showing larger variation is J2228+2201,
although the extreme values span a range of 0.48 dex.

Finally, also the velocity offsets of the \civ\ line are also
approximately constant, and compatible with a single value (within
3$\sigma$).  The exception is J2228+2201, where a significant change
($\sim$7$\sigma$) is observed between the second and third observation
epochs.

The temporal evolution of the velocity offsets, along with the
1450\AA\ continuum luminosity and slope, the \civ\ line FWHM
and virial black hole mass are given in Figure~\ref{fig:QSFit-results}.
